---
title: "Weather Scraper"
author: "Michael Mazel"
date: "1/30/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(XML)
library(dplyr)
library(tidyr)
library(stringi)
library(rvest)
library(RCurl) 
library(lubridate)
library(weathermetrics)
library(openxlsx)
library(dotenv)
```


# Using the MESONET API, we will be scraping weather data by the hour for a given month
Example URL to parse:
https://api.synopticdata.com/v2/stations/timeseries?stid=ktpa&start=202012012353&end=202012012353&vars=air_temp&output=xml&token=your_token_here  
We will insert the MO_DAY and TIME variables using the month_scraper function below.  
Structure of each URL: PRE MO_DAY TIME MID MO_DAY TIME POST TOKEN.   
First, we will create variables to store the PRE MID POST and TOKEN.
The weather variables we will retrieve include air_temp, relative_humidity, and a general weather_summary.
```{r}
PRE <- "https://api.synopticdata.com/v2/stations/timeseries?stid=ktpa&start="
#ktpa is the station ID for Tampa International Airport
MID <- "&end="
POST <- "&vars=air_temp,relative_humidity,weather_summary&output=xml&token="
load_dot_env()
TOKEN = Sys.getenv("token") # My token to use the API is hidden. Insert your token here
```


```{r}
month_scraper <- function(month, year){
  # create a base df to add observations 
  df <- data.frame(date_time=character(),air_temp=integer(),relative_humidity=integer(),weather_summary=character(),stringsAsFactors=FALSE)
  
  # set default days in a month to 31 days
  # create a list of months with 30 days and update the appropriate months
  # update February for leap/non-leap years
  days_in_month <-31
  days_30 <- c(4,6,9,11)
  days_28_29 <- 2
  if (is.element(month, days_30)){
    days_in_month <- 30
  }
  if (is.element(month, days_28_29) & (year %% 4 != 0)){
    days_in_month <- 28
  }
  if (is.element(month, days_28_29) & (year %% 4 == 0)){
    days_in_month <- 29
  }
  
  # make every month a 2 digit string by adding a 0 if necessary
  # adjust "from=1" if you want to start scraping on a day other than the first of the month
  month <- toString(month)
  if (stri_length(month) == 1) {
    month <- paste("0",month,sep="")
    }
  for (day in seq(from=1, to=days_in_month)){
  day <- toString(day)
  
  # make every day a 2 digit string by adding a 0 if necessary
  if (stri_length(day) == 1) {
  day <- paste("0",day,sep="")}
 
  # adjust "by=" if you would like to scrape by an interval other than hour
  # I chose to scrape for every hour at 53 minutes in, because this API produces results most often and most accurately for this minute.
    for(time in seq(from=0053, to=2359, by=100)){
      time <- toString(time)
      # make all times 4 digits
      if (stri_length(time) == 1) {
        time <- paste("000",time,sep="")}
      if (stri_length(time) == 2) {
        time <- paste("00",time,sep="")}
      if (stri_length(time) == 3) {
        time <- paste("0",time,sep="")}
      
      # now we will filter for only numbers that are real times (military time). E.g. 1230, 0245, 1800, 2359
      if ((((substr(time, 1, 1)  == "0") | (substr(time, 1, 1)  == "1")) | ((substr(time, 1, 1)  == "2") & (substr(time, 2, 2) < "4"))) & (substr(time, 3, 3)  < "6")){
        
        # create the URL based off the for-loop "month" and "day" values
        mo_day <- paste(month, day, sep = "")
        combined <-(paste(PRE,year, mo_day,time,MID,year,mo_day,time,POST,TOKEN,sep = ""))
        link_tool <- getURL(combined)
        # To debug function, use: print(combined)
        
        # scrape data from the URL
        doc <- xmlParse(link_tool)
        date_time <- xpathSApply(doc,"//OBSERVATIONS//date_time",xmlValue)
        air_temp <- xpathSApply(doc,"//OBSERVATIONS//air_temp_set_1",xmlValue)
        relative_humidity <- xpathSApply(doc,"//OBSERVATIONS//relative_humidity_set_1",xmlValue)
        weather_summary <- xpathSApply(doc,"//OBSERVATIONS//weather_summary_set_1d",xmlValue)
        
        # enter observation into the df
        observation <- c(date_time,air_temp,relative_humidity,weather_summary)
        df <- rbind(df, observation)
      }
    }
  }
  # rename columns
df <- df %>% rename(date_time = 1, temp = 2, humidity = 3, weather_summary = 4)
  return(df)
}
```

Use the month_scraper function for Feb 2021.
```{r}
Feb21 <- month_scraper(2,2021)
weather <- Feb21
head(weather)
```

Example if scraping multiple months:
```{r}
# Feb21 <- month_scraper(2,2021)
# Mar21 <- month_scraper(3,2021)
# weather <- rbind(Feb21,Mar21)
```

Now, we will clean up our data frame by changing the time zone, converting the temp to F, and filtering out minutes and seconds from date_time.
```{r}
# reformat date_time so we can convert to America/New_York time
substr(weather$date_time, 11, 11) <- " "
weather$date_time = substr(weather$date_time,1,nchar(weather$date_time)-1)
weather$date_time <- force_tzs(ymd_hms(weather$date_time), tzones = "UTC", tzone_out = "America/New_York") # due to the time conversion, a few hours at the end of the month will not be scraped. We will also scrape a few hours from the prior month

# Convert temp to fahrenheit
weather$temp <- as.numeric(weather$temp)
weather$temp <- celsius.to.fahrenheit(weather$temp, round = 1)

# We are only interested in the hour, so we will remove minutes and seconds from date_time
weather$date_time <- as.character(weather$date_time)
weather$date_time <- substr(weather$date_time,1,nchar(weather$date_time)-6)

head(weather)
```

Let's see unique values from the weather summary column.
```{r}
weather$weather_summary %>% unique()
```
The weather_summary variable includes information about cloud coverage and precipitation.  We will create a new column specifically for cloud coverage. Some terms like clear, thin scattered, overcast directly identify the percent of the sky covered. We will convert these to its percent value. Other terms include precipitation info, such as mist or heavy rain. These also have an approximate cloud coverage percent we can use. I researched the approximate percentages.
```{r}
weather <- weather %>% mutate(cloud_cover = weather$weather_summary)

weather$cloud_cover[weather$cloud_cover == "clear"] <- 0
weather$cloud_cover[weather$cloud_cover == "thin scattered"] <- 20
weather$cloud_cover[weather$cloud_cover == "scattered"] <- 45
weather$cloud_cover[weather$cloud_cover == "broken"] <- 75
weather$cloud_cover[weather$cloud_cover == "overcast"] <- 100

# Below are the cloud coverage estimates
weather$cloud_cover[weather$cloud_cover == "light drizzle"] <- 75
weather$cloud_cover[weather$cloud_cover == "light rain"] <- 75
weather$cloud_cover[weather$cloud_cover == "light rain,mist"] <- 100
weather$cloud_cover[weather$cloud_cover == "mist"] <- 100
weather$cloud_cover[weather$cloud_cover == "rain,mist"] <- 100
weather$cloud_cover[weather$cloud_cover == "heavy rain,mist"] <- 100
weather$cloud_cover[weather$cloud_cover == "fog"] <- 75
weather$cloud_cover[weather$cloud_cover == "haze"] <- 45
# Note, this is not a complete list based off all the above unique weather_summary categories

weather$cloud_cover <- as.numeric(weather$cloud_cover)
```

We will create a new column for precipitation. We will convert to the approximate precipitation rate in mm/hr. This may not be exactly what was measured from the API, but it should be close.
```{r}
weather$precip[weather$weather_summary == "light drizzle"] <- 1
weather$precip[weather$weather_summary == "light rain"] <- 1.25
weather$precip[weather$weather_summary == "light rain,mist"] <- 1.5
weather$precip[weather$weather_summary == "mist"] <- 4.5
weather$precip[weather$weather_summary == "rain,mist"] <- 5
weather$precip[weather$weather_summary == "heavy rain,mist"] <- 10
weather$precip[weather$weather_summary == "fog"] <- 0
weather$precip[weather$weather_summary == "haze"] <- 0
weather$precip[weather$weather_summary == "clear"] <- 0
weather$precip[weather$weather_summary == "thin scattered"] <- 0
weather$precip[weather$weather_summary == "scattered"] <- 0
weather$precip[weather$weather_summary == "broken"] <- 0
weather$precip[weather$weather_summary == "overcast"] <- 0

```

Create a heat index column. This is calculated from the air temperature and humidity.
```{r}
weather$humidity <- as.numeric(weather$humidity)
weather <- weather %>% 
mutate(heat_index = heat.index(t = weather$temp,rh = weather$humidity, temperature.metric = 'fahrenheit', output.metric = 'fahrenheit', round = 1))
```

```{r}
head(weather,24)
```
